### 2017-04-11 progress/update.tex
add update, small mdl

Tumors are composed of different sub-populations of cells, and these sub-populations exhibit shared patterns of gene expression. With contemporary sequencing machines, it is possible to obtain the expression levels of 10,000+ genes for 1000+ cells in a single experiment. In this project, we aim to apply Bayesian methods for clustering and topic modeling to discover meaningful cell clusters and gene modules from single-cell RNA-sequencing (scRNA-seq) datasets from tumors. We first intend to explore the use of a hierarchical topic model, Latent Dirichlet Allocation (LDA) \cite{LDA}, as well as two non-parametric topic models, Hierarchical Dirichlet Processes (HDP) \cite{HDP} and the Indian Buffet Process Compound Dirichlet Process (IBP-CDP) \cite{IBP} for analyzing scRNA-seq data. We will evaluate these models based on their ability to discover meaningful functional gene modules. Subsequently, we plan to explore the use of mixture models to cluster cells based on the output of the topic models, which we will evaluate based on their ability to differentiate cell types (e.g. immune, cancer, non-cancerous). Finally, we will train a combined clustering-topic model (e.g. MGCTM \cite{pengtao}) to see if it outperforms the individual models. We will compare the results of these methods with standard methods used in prior work, and evaluate these methods' robustness to noise. Time permitting, we will study extensions of these models appropriate for time-series scRNA-seq data \cite{TS-DDP}. \\

Prior research has applied basic measures of statistical distance to cluster cell groups or gene modules based on samples of single-cell RNA-sequencing (scRNA-seq) data taken from tumors \cite{nature}.  Good clustering of scRNA-seq data often has biologically meaningful results, and as such, a wide variety of correlation based methods have been used in prior work to derive meaning from scRNA-seq data \cite{coexpression, consensus, profiling}. We believe that Bayesian methods for clustering and topic modeling can be more illuminating than these previous methods, due to their consideration of higher-order relationships within the data. \\

In summary, and in order of importance, we are primarily interested in answering the following questions:

>>>>

The progress report (one per team) will be a written pdf document, about 2 pages long,
reporting on what has been accomplished so far and outlining the work that remains to be done.
You should write clearly about how you’ve spent your time since the proposal and any preliminary
findings (positive or negative). If there are multiple participants, the division of time spent
so far should be made clear. Then you should include a plan (updated and/or revised from the
proposal, as appropriate) for the time remaining. If there are multiple participants, the division of
responsibility going forward should be made clear.
